### 2021-04-13 submission/manuscript.Rmd
Update manuscript with latest figures
Add to results section

**Laxative treatment alone leads to prolonged *C. difficile* colonization in mice.** We compared PEG-treated mice to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing C. difficile within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). After PEG and/or antibiotic treatment all mice were challenged with 10^3^ *C. difficile* 630 spores.

@@

* Figure 5. 1-day PEG treatment post C. difficile challenge prolongs colonization regardless of whether an FMT is also administered.

@@

\includegraphics{figure_5.pdf}
\includegraphics{figure_5_16S.pdf}
**Figure 5. 1-day PEG treatment post C. difficile challenge prolongs colonization regardless of whether an FMT is also administered.** A. 
\newpage

>>>>

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared PEG-treated mice to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^3^ *C. difficile* 630 spores. All treatments rendered mice susceptible to *C. difficile* colonization (Fig. 1C), however PEG-treated mice remained colonized at a high level through 30 days post-infection. In contrast, the clindamycin-treated mice that cleared *C. difficile* within 10 days post-infection. Surprisingly, mice were still susceptible to *C. difficile* infection after 10-days of recovery from PEG treatment although *C. difficile* was not detectable in most of the group in the initial 5 days post-infection (Fig. 1C). From 9 days post-infection onward, the median *C. difficile* stabilized for the 5-day PEG plus 10-day recovery group of mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible for up to 10 days post PEG treatment.
**5-day laxative treatment differentially disrupts the fecal microbiota compared to clindamycin treatment.** Given 
**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given

@@

* Figure 5. 1-day PEG treatment post *C. difficile* challenge prolongs colonization regardless of whether an FMT is also administered.

@@

\includegraphics{figure_5.pdf}
\newpage
\includegraphics{figure_5_16S.pdf}
**Figure 5. 1-day PEG treatment post C. difficile challenge prolongs colonization regardless of whether an FMT is also administered.** A. 
\newpage

### 2021-04-13 submission/manuscript.tex
Update manuscript with latest figures
Add to results section

\textbf{Laxative treatment alone leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing C.
difficile within 10 days post-infection (9, 19). All PEG-treated mice
were administered a 15\% PEG solution in the drinking water for 5-days,
one group was also treated with clindamycin, and one group was allowed
to recover for 10 days prior to challenge (Fig. 1A). After PEG and/or
antibiotic treatment all mice were challenged with 10\textsuperscript{3}
\emph{C. difficile} 630 spores.

\begin{itemize}
\item
  Figure 1. 5-day PEG treatment prolongs susceptibility and mice become
  persistently colonized with \emph{C. difficile}.
\item
  Figure 2. 5-day PEG treatment disrupts the stool microbiota for a
  longer amount of time compared to clindamycin-treated mice.
\item
  Figure S1. 5-day PEG treatment plus 10-day recovery mice microbiota
  dynamics post-infection.
\item
  Figure 3. 5-day PEG treatment does not result in more severe CDIs,
  although mucosal microbiota is altered.
\item
  Figure 4. 1-day PEG treatment renders mice susceptible to transient
  \emph{C. difficile} colonization.
\item
  Figure 5. 1-day PEG treatment post C. difficile challenge prolongs
  colonization regardless of whether an FMT is also administered.
\item
  Figure 6. Specific microbiota features associated with prolonged
  \emph{C. difficile} colonization in PEG treated mice.
\item
  Figure S2. Specific OTUs associated with clearance that are mostly
  absent in mice with prolonged \emph{C. difficile} colonization.

@@

\includegraphics{figure_5.pdf} \includegraphics{figure_5_16S.pdf}
\textbf{Figure 5. 1-day PEG treatment post C. difficile challenge
prolongs colonization regardless of whether an FMT is also
administered.} A. \newpage

>>>>

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing
\emph{C. difficile} within 10 days post-infection (9, 19). All
PEG-treated mice were administered a 15\% PEG solution in the drinking
water for 5-days, one group was also treated with clindamycin, and one
group was allowed to recover for 10 days prior to challenge (Fig. 1A).
PEG treatment resulted in weight loss in all 3 groups of PEG-treated
mice, with the greatest change in weight observed on the fifth day of
PEG treatment (Fig. 1B). After either PEG, clindamycin, or PEG and
clindamycin treatment all mice were challenged with
10\textsuperscript{3} \emph{C. difficile} 630 spores. All treatments
rendered mice susceptible to \emph{C. difficile} colonization (Fig. 1C),
however PEG-treated mice remained colonized at a high level through 30
days post-infection. In contrast, the clindamycin-treated mice that
cleared \emph{C. difficile} within 10 days post-infection. Surprisingly,
mice were still susceptible to \emph{C. difficile} infection after
10-days of recovery from PEG treatment although \emph{C. difficile} was
not detectable in most of the group in the initial 5 days post-infection
(Fig. 1C). From 9 days post-infection onward, the median \emph{C.
difficile} stabilized for the 5-day PEG plus 10-day recovery group of
mice and remained high through 30 days post-infection (Fig. 1C). Thus,
osmotic laxative treatment alone was sufficient to render mice
susceptible to prolonged \emph{C. difficile} colonization and
PEG-treated mice remained susceptible for up to 10 days post PEG
treatment. \textbf{5-day laxative treatment differentially disrupts the
fecal microbiota compared to clindamycin treatment.} Given \textbf{5-day
laxative treatment does not promote more severe CDIs despite altering
the mucosal microbiota.} Given

\begin{itemize}
\item
  Figure 1. 5-day PEG treatment prolongs susceptibility and mice become
  persistently colonized with \emph{C. difficile}.
\item
  Figure 2. 5-day PEG treatment disrupts the stool microbiota for a
  longer amount of time compared to clindamycin-treated mice.
\item
  Figure S1. 5-day PEG treatment plus 10-day recovery mice microbiota
  dynamics post-infection.
\item
  Figure 3. 5-day PEG treatment does not result in more severe CDIs,
  although mucosal microbiota is altered.
\item
  Figure 4. 1-day PEG treatment renders mice susceptible to transient
  \emph{C. difficile} colonization.
\item
  Figure 5. 1-day PEG treatment post \emph{C. difficile} challenge
  prolongs colonization regardless of whether an FMT is also
  administered.
\item
  Figure 6. Specific microbiota features associated with prolonged
  \emph{C. difficile} colonization in PEG treated mice.
\item
  Figure S2. Specific OTUs associated with clearance that are mostly
  absent in mice with prolonged \emph{C. difficile} colonization.

@@

\includegraphics{figure_5.pdf} \newpage
\includegraphics{figure_5_16S.pdf} \textbf{Figure 5. 1-day PEG treatment
post C. difficile challenge prolongs colonization regardless of whether
an FMT is also administered.} A. \newpage
